The five axes of the Turtle: symmetry and asymmetry in NGC~6210

### 2020-11-29 ms.tex
Refer to the new Fig 4 (1D spectra) in more places

@@

All data was reduced using standard IRAF\footnote{IRAF was originally
  distributed by the National Optical Astronomy Observatory,
  \url{http://ast.noao.edu/data/software},
  and is currently available from \url{https://iraf-community.github.io}.}
routines.
Bias and cosmic-ray removal was followed by rectification and first-order wavelength 
calibration of the two-dimensional spectra
based on the comparison spectrum
of a Th/Ar lamp, yielding an accuracy in velocity units of \SI{\pm 1}{km.s^{-1}}.
Figure~\ref{fig:pv-array} shows example position-velocity arrays of (left-to-right)
\Ha{} \Wav{6563}, \oiii{} \Wav{5007}, and \nii{} \Wav{6583} for the 11 individual positions from the 2015 run.
The velocity scale is heliocentric 
and the scale along the slit is specified as arcsecond offsets from the central star.
% We derive a heliocentric
% systemic velocity, \(\vsys = \SI{40}{km.s^{-1}}\) by using the slit position F (epoch 2003) that passes through the central star and is not saturated. 
An additional slit (x, shown in Figure~\ref{fig:halo-knots}) was placed on the northern outskirts of the nebula 
in the outer halo of NGC~6210 where filaments had been clearly detected in some of our deep images.\footnote{For this spectrum only, \(3 \times 3\) pixel on-chip binning was used.}
% The observation was carried out on 18 September 2019.  The spectrum was obtained in order 87 (\Ha{} and \nii lines) with the 150 micron slit oriented at a position angle of 56 degrees.  The detector was again the Marconi 2048x2048 with a 3$\times$3 binning.  The spectral dispersion was 0.087 \AA/pix and 0.525\arcsec/pixel.
Most of the spectra presented here are publicly available from the \mbox{San Pedro Mártir} Kinematic Catalogue of Galactic Planetary Nebulae\footnote{
  \url{http://kincatpn.astrosen.unam.mx}
}
\citep{Lopez:2012a}.

@@

In order to investigate the kinematics of the nebula in detail,
we have measured the velocities of distinct emission components in each slit spectrum
and organized them into broad systems based on their location, morphology and degree of ionization.
Maps of the individual slit positions,
marked with the position and velocity of each component
are presented in Figures~\ref{fig:shell-velocity-components} to \ref{fig:halo-components}.
All line-of-sight velocities are given with respect to a nominal systemic heliocentric velocity of \SI{-40}{km.s^{-1}},
which we choose as a typical velocity of the planetary nebula halo. 
The figures are organized according to the different nebular features shown in Figure~\ref{fig:hst}:
shells (Figure~\ref{fig:shell-velocity-components} for \oiii{} and Figure~\ref{fig:heii-shell-components} for \heii{}),
knot complexes (Figure~\ref{fig:knot-complex-map}),
outer lobes (Figure~\ref{fig:outer-lobe-components}),
and halo (Figure~\ref{fig:halo-components}).
Each of these features is discussed in more detail below.

@@

The knot complexes are also visible in the \oiii{} spectra,
although they are fainter than the shells in this line.
There is also a slight difference in velocity,
with \oiii{} showing expansion velocities that are \num{2} to \SI{5}{km.s^{-1}} smaller than \nii{}.
If the expansion velocities within the knot complexes are assumed to increase with radius,
then this is evidence for ionization stratification along the line of sight
with the \oiii{} emission arising slightly closer to the star than \nii{}.
This is also consistent with the HST images (e.g., Figure~\ref{fig:hst}d),
which show that the \oiii{} emission associated with the complexes seems to arise from the inner surface of the\nii{} knots.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/turtle-1d-line-profiles}
  \caption{
    Thick colored lines show one-dimensional extracted velocity profiles
    of (top to bottom) \nii{} (red), \Ha{} (purple) and \oiii{} (light blue)
    for a  short section of slit~e
    (length \SI{2}{arcsec},
    centered \SI{3}{arcsec} north of the central star).
    For the \nii{} and \oiii{} lines, four-gaussian fits are shown
    (continuous thin black lines show the individual components,
    while thin dashed line shows the total profile).
    The bottom panel shows the residuals of these fits on an expanded scale.
    Short vertical colored lines mark the centroid velocity of each of the four components,
    with thicker lines corresponding to the high-ionization shell
    and thinner lines low-ionization knot complexes.
    The vertical dashed line shows the systemic velocity.
    No fits are performed to the \Ha{} line
    because the larger thermal broadening means that
    the four components cannot be reliably separated.
  }
  \label{fig:spec-1d}
\end{figure}

@@

All data was reduced using standard IRAF\footnote{IRAF was originally
  distributed by the National Optical Astronomy Observatory,
  \url{http://ast.noao.edu/data/software},
  and is currently available from \url{https://iraf-community.github.io}.}
routines.
Bias and cosmic-ray removal was followed by rectification and first-order wavelength 
calibration of the two-dimensional spectra
based on the comparison spectrum
of a Th/Ar lamp, yielding an accuracy in velocity units of \SI{\pm 1}{km.s^{-1}}.
Figure~\ref{fig:pv-array} shows example position-velocity arrays of (left-to-right)
\Ha{} \Wav{6563}, \oiii{} \Wav{5007}, and \nii{} \Wav{6583} for the 11 individual positions from the 2015 run.
The velocity scale is heliocentric 
and the scale along the slit is specified as arcsecond offsets from the central star.
Figure~\ref{fig:spec-1d} shows example one-dimensional spectra,
extracted from a short section of slit~e.
% We derive a heliocentric
% systemic velocity, \(\vsys = \SI{40}{km.s^{-1}}\) by using the slit position F (epoch 2003) that passes through the central star and is not saturated. 
An additional slit (x, shown in Figure~\ref{fig:halo-knots}) was placed on the northern outskirts of the nebula 
in the outer halo of NGC~6210 where filaments had been clearly detected in some of our deep images.\footnote{For this spectrum only, \(3 \times 3\) pixel on-chip binning was used.}
% The observation was carried out on 18 September 2019.  The spectrum was obtained in order 87 (\Ha{} and \nii lines) with the 150 micron slit oriented at a position angle of 56 degrees.  The detector was again the Marconi 2048x2048 with a 3$\times$3 binning.  The spectral dispersion was 0.087 \AA/pix and 0.525\arcsec/pixel.
Most of the spectra presented here are publicly available from the \mbox{San Pedro Mártir} Kinematic Catalogue of Galactic Planetary Nebulae\footnote{
  \url{http://kincatpn.astrosen.unam.mx}
}
\citep{Lopez:2012a}.

@@

In order to investigate the kinematics of the nebula in detail,
we have measured the velocities of distinct emission components in each slit spectrum
and organized them into broad systems based on their location, morphology and degree of ionization.
The velocities were determined by fitting gaussians to one-dimensional line profiles,
as illustrated in Figure~\ref{fig:spec-1d}.
The higher ionization features are measured in \oiii{},
while the lower ionization features are measured in \nii{}.

Maps of the individual slit positions,
marked with the position and velocity of each component
are presented in Figures~\ref{fig:shell-velocity-components} to \ref{fig:halo-components}.
All line-of-sight velocities are given with respect to a nominal systemic heliocentric velocity of \SI{-40}{km.s^{-1}},
which we choose as a typical velocity of the planetary nebula halo. 
The figures are organized according to the different nebular features shown in Figure~\ref{fig:hst}:
shells (Figure~\ref{fig:shell-velocity-components} for \oiii{} and Figure~\ref{fig:heii-shell-components} for \heii{}),
knot complexes (Figure~\ref{fig:knot-complex-map}),
outer lobes (Figure~\ref{fig:outer-lobe-components}),
and halo (Figure~\ref{fig:halo-components}).
Each of these features is discussed in more detail below.

@@

The knot complexes are also visible in the \oiii{} spectra,
although they are fainter than the shells in this line.
There is also a slight difference in velocity,
with \oiii{} showing expansion velocities that are \num{2} to \SI{5}{km.s^{-1}} smaller than \nii{},
as can be seen in Figure~\ref{fig:spec-1d}.
If the expansion velocities within the knot complexes are assumed to increase with radius,
then this is evidence for ionization stratification along the line of sight
with the \oiii{} emission arising slightly closer to the star than \nii{}.
This is also consistent with the HST images (e.g., Figure~\ref{fig:hst}d),
which show that the \oiii{} emission associated with the complexes seems to arise from the inner surface of the\nii{} knots.
